---
title: "SQL"
author: "Micheal Rekitt"
date: "October 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#SQL

## Filtering Rows and Columns
SQL can be used to filter columns and rows. SELECT is used to filter columns and WHERE is used to filter rows. Using the Lahman and sqldf packages we can do some examples.

#### Required Packages
```{r}
library(Lahman)
library(sqldf)
```
#### Example
For this example we will look at the homerun totals for the 1927 Yankees.

```{r}
query<-"SELECT playerID, teamID, yearID, HR 
FROM Batting
WHERE teamID='NYA' and yearID=1927"

sqldf(query)
```

## Order By
SQL also has the ability to order the rows. As an example, adding ORDER BY HR DESC will order the results by Homeruns in descending order.

#### Example
```{r}
query<-"SELECT playerID, teamID, yearID, HR 
FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```
You can ORDER BY any factor.

#### Example
```{r}
query<-"SELECT playerID, yearID, teamID, SO, AB 
FROM Batting 
WHERE SO<10 and AB>400 
ORDER BY SO" 

sqldf(query) 
```
##Aggregation

Aggregation allows you to group the data. Using GROUP BY will perform the command. There a multiple ways to Group the data:

*sum(colname) gives you the sum

*avg(colname) gives you the average

*max(colname) gives you the maximum

*min(colname) gives you the minimum

#### Examples
##### #1
```{r}
query<-"SELECT playerID, sum(HR) 
FROM Batting 
WHERE playerID='ruthba01' 
GROUP BY playerID" 

sqldf(query) 
```
##### #2
```{r}
query<-"SELECT playerID, sum(HR)
FROM Batting 
GROUP BY playerID 
HAVING sum(HR)>=600 
ORDER BY sum(HR) DESC" 

sqldf(query) 
```
##### #3
```{r}
query<-"SELECT playerID, avg(HR)
FROM Batting 
GROUP BY playerID 
HAVING avg(HR)>30 
ORDER BY avg(HR) DESC" 

sqldf(query)
```
## Joining
Joining allows data from multiple tables to be used together by using the commands INNER JOIN and ON. 

#### Examples
##### #1
```{r}
query<-"SELECT nameFIRST, nameLAST, teamID, yearID, HR  
FROM Batting INNER JOIN Master 
ON Batting.playerID=Master.playerID 
WHERE Batting.playerID='ruthba01'" 

sqldf(query) 
```
##### #2
```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR 
FROM Batting INNER JOIN Master 
ON Batting.playerID=Master.playerID 
WHERE Batting.HR>50
ORDER BY HR DESC" 

sqldf(query) 
```
##### #3
```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR 
FROM Batting INNER JOIN Teams 
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID 
WHERE playerID='ruthba01'" 

sqldf(query) 
```

